Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework

### 2015-04-09 tex/manuscript.tex
Worked in most comments

%% ----------------------------------
%
%     Packages to be used for drafts and editing
%     Remove from the final manuscript
%
%% ----------------------------------
%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}

@@

\title{Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework}
\date{}
\author[1,2,13]{Matthew V. Talluto (mtalluto@gmail.com)}
\author[1,2]{Isabelle Boulangeat (isabelle.boulangeat@gmail.com)}
\author[3]{Aitor Ameztegui (ameztegui@gmail.com)}
\author[4]{Isabelle Aubin (Isabelle.Aubin@NRCan-RNCan.gc.ca)}
\author[1,2,5]{Dominique Berteaux (Dominique\_Berteaux@uqar.ca)}
\author[1,2]{Alyssa Butler (ca.butler10@gmail.com)}
\author[6,7]{Frédérik Doyon (Frederik.Doyon@uqo.ca)}
\author[8]{C. Ronnie Drever (cdrever@TNC.ORG)}
\author[9]{Marie-Josée Fortin (fortinmj@gmail.com)}
\author[1]{Tony Franceschini (Tony.Franceschini@uqar.ca)}
\author[10]{Jean Liénard (jean.lienard@gmail.com)}
\author[4]{Dan McKenney (Dan.McKenney@nrcan-rncan.gc.ca)}
\author[2,3]{Kevin A. Solarik (kevinsolarik@hotmail.com)}
\author[10]{Nikolay Strigul (nick.strigul@vancouver.wsu.edu)}
\author[11, 12]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
\author[1,2]{Dominique Gravel (dominique\_gravel@uqar.ca)}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[6]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
\affil[7]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
\affil[8]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
\affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
\affil[10]{Department of Mathematics, Washington State University, Vancouver, Washington, USA}
\affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[13]{Author for correspondance. Address: Departament de Biologie, chimie, et geographie, 300, Allée des Ursulines, Rimouski, Quebec G5L 3A1, Canada}

@@

	\textbf{Abstract word count: } 258 %ABS_COUNT

	\textbf{Main text word count: } 4641 %WORD_COUNT %Target 5000

	\textbf{Authorship:} All authors contributed substantially to an initial outline of the paper and ideas. MT, IB, and DG built the model and wrote the first draft of the introduction and model description. All authors contributed to the complete first draft and all subsequent revisions to the manuscript.

	\textbf{Keywords:} Climate change, decision making, metamodeling, patterns and processes, range dynamics, scaling, spatial ecology, species distribution modeling, uncertainty
	\end{flushleft}
\end{titlingpage}
%TC:endignore

@@

Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
However, in general, individual approaches include only a small subset of the available information about a species. 
Correlative models often predict presence or absence as a function of climate, but ignore  smaller-scale processes such as growth, fecundity, and dispersal.
Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
Our framework uses a metamodel, estimated in a hierarchical Bayesian framework, that is constrained by both data directly applied to the metamodel and by the results of one or more sub-models.
The output of the metamodel is in the form of probabilistic estimates of species presence.
Metamodel outputs also incorporate uncertainty from both data and process error from all sub-models, and include all of the information used as input for the original scale-specific sub-models.
We illustrate our approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of the original data.
We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-scale data into ecological decision-making.
Although implementation of our method can be technically challenging, we anticipate the results will be of wide interest, and we encourage collaboration between modelers and practitioners in applying our framework.

@@

Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}. 
Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Blois2013, Thuiller2014}.

Despite the recognized potential of these models, it can be difficult to produce species distribution models (SDMs) with acceptable levels of precision and bias \citep{Guisan2013}.
For mechanistic models, two important constraints can present difficulties: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
In recent decades, however, modeling techniques have proliferated to take advantage of increased data availability. 
A growing body of theory, reflecting the diversity of processes generating species ranges, also has contributed to model diversification \citep{Boulangeat2012}.
Of these model types, fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, perform poorly when applied at the scale of species ranges.
For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Pigot2013}. 
In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}. 
Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
%% ----------------------------------
%% ----------------------------------
%% ----------------------------------

Clearly, an approach is needed that unifies the strength of the diversity of modeling approaches and can incorporate multiple data sources and modeling paradigms.
Toward this end, we present here an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
Techniques for multimodel inference have proliferated in recent years.
For example, hybrid models that allow for flexible combinations between mechanistic and phenomenological sub-models are commonly employed in SDMs \citep{Gallien2010, Thuiller2013, Boulangeat2014}.
Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include space and time dynamics and therefore account for dispersal \citep{Smolik2010}.
However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and the sources and magnitudes of uncertainties are difficult to quantify. 
A commonly-employed alternative to hybrid models is the direct combination of predictions, where predictions from models operating at the same spatio-temporal scales can be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}. 
However, the applicability of such techniques is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.

We propose an alternative to these other approaches to multimodel inference.
We use the power and generality of a hierarchical Bayesian framework, which provides a number of desirable advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) intuitive and comprehensive reporting of uncertainty in model predictions that  reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}.
Contrary to hybrid and ensemble methods, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included. 
By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Guisan2005, Araujo2006}.
Moreover, a more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.

We illustrate our approach with two examples.
We begin with a hypothetical example, using simulated data, where we define the framework and demonstrate its application to multiple sources of information from different scales.
In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
Both examples are presented with complete code and data (in Supporting Information).
Finally, we provide a more formal mathematical presentation in Appendix S1.

%% ----------------------------------
%
%     EXAMPLE 1
%
%% ----------------------------------
\section*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions, and to constrain the parameters of this model using the output of one or more sub-models
In this hypothetical example, we build a metamodel relating the distribution (i.e. occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species. 
However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
Thus, we use additional information on the physiological constraints affecting species distribution.
Because these data are often collected at a finer scale than that of rangewide occurence data, we apply a simple scaling function, drawing on ecological theory, to compute the likelihood of a set of metamodel parameters given both the occurrence and the physiological data.
See Appendix S1 in Supporting Information for procedural details and scripts for executing the model.

%==================
%TC:ignore
\begin{figure}[t]
	\includegraphics{figs/ex1_sampling.pdf}
	\caption{Two simulated datasets used to illustrate the model integration framework.
	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
	\label{fig:ex1_sampling}
\end{figure}
%TC:endignore
%==================

We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation.
We assume a presence-absence dataset, which we term \(X_M\), where the subscripted \(M\) indicates that the data were collected at the same scale as the metamodel.
We simulated a relatively high-quality dataset that provides coverage of a variety of ecological conditions (Fig. \ref{fig:ex1_sampling}).
We desire to model the species' distribution as a function of temperature, \(T_M\), and precipitation, \(P_M\).
An initial version of the metamodel \((\theta_M)\) that has no constraints from other datasets will be referred to as the naive model.
This naive model uses a simple Bayesian logistic regression to estimate the naive probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
%-----
\begin{equation}
\begin{aligned}
	\psi_N &= f\left(\theta_M, T_M, P_M \right) \\
	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
	&=\text{logit}^{-1}\left( \theta_M D_M \right)
\end{aligned}
\end{equation}
%-----
where \(theta_M\) is the parameter vector of the model, \(D_M\) is the covariate matrix (i.e., \(T_M, P_M\), and quadratic terms, with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
We estimate parameters using a Markov Chain Monte Carlo (MCMC) method using the proportional form of Bayes' theorem:
%-----
\begin{equation}
\label{eq:ex1_bayes}
	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
	p \left(X_M \mid \theta_M, T_M, P_M \right)
	p \left(\theta_M \right)
\end{equation}
%-----
where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.

Thus far, we have considered only a single source of information to fit this model, and therefore the prior distribution \(p\left(\theta_M \right)\) from Eq. \ref{eq:ex1_bayes} is uninformative.
As a secondary source of information, we will consider an experiment relating the population growth rate of the plant to manipulations to the precipitation regime, with results (but no raw data) available from the literature (Fig. \ref{fig:ex1_sampling}b). 
Furthermore, no information is available regarding the temperature regime for the experiment.
Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
This concept gives us a reasonable model to fit a scaling function for our sub-model (Appendix S1).
If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for an observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
\begin{equation}
	\psi_{S,i} = \int_0^\infty N \left(X_{S,i}, \sigma_{S,i} \right)
\end{equation}
where \(N\) is the Normal density function.
We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
%-----
\begin{equation}
\label{eq:ex1_thetas}
	p\left (\theta_S \mid \psi_S,P_S \right ) \propto 
	p \left(\psi_S \mid \theta_S, P_S \right)
	p \left(\theta_S \right) \\
\end{equation}
%-----

Although the two datasets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation). 
The scaling is quite simplistic; it treats the fundamental niche as the only driver of species distribution, and only considers a single dimension of the niche.
As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
As we will show, however, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.

>>>>

%% ----------------------------------
%
%     Packages to be used for drafts and editing
%     Remove from the final manuscript
%
%% ----------------------------------
\usepackage[top=1in, left=1in, bottom=1in, right=2.0in]{geometry}
\usepackage[textsize=tiny, backgroundcolor=white, textwidth=1.8in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
\newcommand{\ib}[1]{\todo[color=red!40]{\textsuperscript{IB}#1}}
\newcommand{\fd}[1]{\todo[color=orange!40]{\textsuperscript{FD}#1}}

@@

\title{Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework}
\date{}
\author[1,2,13]{Matthew V. Talluto (mtalluto@gmail.com)}
\author[1,2]{Isabelle Boulangeat (isabelle.boulangeat@gmail.com)}
\author[3]{Aitor Ameztegui (ameztegui@gmail.com)}
\author[4]{Isabelle Aubin (Isabelle.Aubin@NRCan-RNCan.gc.ca)}
\author[1,2,5]{Dominique Berteaux (Dominique\_Berteaux@uqar.ca)}
\author[1,2]{Alyssa Butler (ca.butler10@gmail.com)}
\author[6,7]{Frédérik Doyon (Frederik.Doyon@uqo.ca)}
\author[8]{C. Ronnie Drever (cdrever@TNC.ORG)}
\author[9]{Marie-Josée Fortin (mariejosee.fortin@utoronto.ca)}
\author[1]{Tony Franceschini (Tony.Franceschini@uqar.ca)}
\author[10]{Jean Liénard (jean.lienard@gmail.com)}
\author[4]{Dan McKenney (Dan.McKenney@nrcan-rncan.gc.ca)}
\author[2,3]{Kevin A. Solarik (kevinsolarik@hotmail.com)}
\author[10]{Nikolay Strigul (nick.strigul@vancouver.wsu.edu)}
\author[11, 12]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
\author[1,2]{Dominique Gravel (dominique\_gravel@uqar.ca)}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[6]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
\affil[7]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
\affil[8]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
\affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
\affil[10]{Department of Mathematics, Washington State University, Vancouver, Washington, USA}
\affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[13]{Author for correspondance. Address: Departament de Biologie, chimie, et geographie, 300, Allée des Ursulines, Rimouski, Quebec G5L 3A1, Canada}

@@

	\textbf{Abstract word count: } 235 %ABS_COUNT

	\textbf{Main text word count: } 4848 %WORD_COUNT %Target 5000

%	\textbf{Authorship:} All authors contributed substantially to an initial outline of the paper and ideas. MT, IB, and DG built the model and wrote the first draft of the introduction and model description. All authors contributed to the complete first draft and all subsequent revisions to the manuscript.

	\textbf{Keywords:} Climate change, decision making, patterns and processes, range dynamics, scaling, spatial ecology, species distribution modeling, uncertainty
	\end{flushleft}
\end{titlingpage}
%TC:endignore

@@

Current interest in modeling potential changes to species ranges have resulted in a great diversity of approaches to species distribution modeling.
However, in general, single approaches include only a small subset of the available information about a species. 
Correlative models often predict presence or absence as a function of the environment, but ignore smaller-scale processes such as growth, fecundity, and dispersal.
Furthermore, different approaches often produce divergent predictions, with no simple method to reconcile them.
We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
Our framework uses a metamodel, estimated in a hierarchical Bayesian framework, that is constrained by both data directly applied to the metamodel and by the results of one or more sub-models.
The output of the metamodel is in the form of probabilistic estimates of species presence.
Metamodel outputs also incorporate uncertainty from both data and process error from all sub-models, and include all of the information used as input for the original scale-specific sub-models.
We illustrate our approach using theoretical and empirical examples, and demonstrate that the framework better characterizes uncertainty and can substantially reduce uncertainty when projecting beyond the range of the original data.
We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
The framework has wide potential for use in species distribution modelling, and can drive better integration of multi-scale data into ecological decision-making.

@@

Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}. 
Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, hybrid zone dynamics \citep{Engler2013}, and climate change impacts on species distributions \citep{Blois2013, Thuiller2014}.

Despite the recognized potential of these models, it can be difficult to produce species distribution models (SDMs) with acceptable levels of precision and bias \citep{Guisan2013}.
For mechanistic models, two important constraints can cause difficulties: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
In recent decades, however, modeling techniques have proliferated to take advantage of increased data availability. 
A growing body of theory, reflecting the diversity of processes generating species ranges, also has contributed to model diversification \citep{Boulangeat2012}.
Of these model types, fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, perform poorly when applied at the scale of species ranges.
For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Pigot2013}. 
In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}. 
Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
%% ----------------------------------
%% ----------------------------------
%% ----------------------------------

Clearly, an approach is needed to unify the strength of different modeling approaches that can incorporate multiple data sources and modeling paradigms.
Toward this end, we present here an application of hierarchical Bayesian methods that uses outputs from multiple models to inform the results of the final model.
Techniques for multimodel inference have proliferated in recent years.
For example, hybrid models that allow for flexible combinations between mechanistic and phenomenological sub-models are commonly employed in SDMs \citep{Gallien2010, Thuiller2013, Boulangeat2014}.
Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include biotic interactions and space-time dynamics through dispersal \citep{Smolik2010}.
However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and uncertainties are approximated and their attribution to different sources is difficult to distinguish.
A commonly-employed alternative to hybrid models is the direct combination of predictions, where predictions from models operating at the same spatio-temporal scales can be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}. 
However, the applicability of such techniques is limited; for example, because uncertainty is approximated and may be poorly understood, it is not possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.

We propose an alternative to these other approaches using a hierarchical Bayesian framework. 
This approach provides a number of desirable advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) an intuitive and comprehensive reporting of uncertainty in model predictions that  reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}.
Contrary to hybrid methods, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for more flexibility regarding the type of information included. 
By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Guisan2005, Araujo2006}.
Moreover, a more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.

We illustrate our framework with two examples.
We begin with a hypothetical example, using simulated data, where we define the framework and demonstrate its application to multiple sources of information from different scales.
In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
We provide a more formal mathematical presentation (Appendix S1) along with complete code and data for both examples (Appendix S2).

%% ----------------------------------
%
%     EXAMPLE 1
%
%% ----------------------------------
\section*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
The key idea of our approach is to formulate a metamodel that integrates data at the same ecological scale as the desired predictions, and to constrain the parameters of this model using the output of one or more sub-models.
In this hypothetical example, we build a metamodel relating the distribution of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species. 
However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
Thus, we use additional information on the physiological constraints affecting species distribution.
Because these data are often collected at a finer scale than that of rangewide occurence data, we apply a simple scaling function, drawing on ecological theory, to compute the likelihood of a set of metamodel parameters given both the occurrence and the physiological data.
See Appendix S1 for procedural details and scripts for executing the model.

%==================
%TC:ignore
\begin{figure}[t]
	\includegraphics{figs/ex1_sampling.pdf}
	\caption{Two simulated datasets used to illustrate the model integration framework.
	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., -1.0--1.0).
	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
	\label{fig:ex1_sampling}
\end{figure}
%TC:endignore
%==================

We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation.
For the metamodel-scale data, we simulate a relatively high-quality presence-absence dataset covering a variety of ecological conditions, which we term \(X_M\), where the subscripted \(M\) indicates that the data were collected at the same scale as the metamodel (Fig. \ref{fig:ex1_sampling}).
We desire to mo0del the species' distribution as a function of temperature, \(T_M\), and precipitation, \(P_M\).
An initial version of the metamodel \((\theta_M)\) that has no constraints from other datasets will be referred to as the naive model.
This naive model uses a simple Bayesian logistic regression to estimate the naive probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
%-----
\begin{equation}
\begin{aligned}
	\psi_N &= f\left(\theta_M, T_M, P_M \right) \\
	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
	&=\text{logit}^{-1}\left( \theta_M D_M \right)
\end{aligned}
\end{equation}
%-----
where \(\theta_M\) is the parameter vector of the model, \(D_M\) is the covariate matrix (i.e., \(T_M, P_M\)), and \(\text{logit}^{-1}\) is the inverse of the logit function.
We estimate parameters using a Markov Chain Monte Carlo (MCMC) method using the proportional form of Bayes' theorem:
%-----
\begin{equation}
\label{eq:ex1_bayes}
	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
	p \left(X_M \mid \theta_M, T_M, P_M \right)
	p \left(\theta_M \right)
\end{equation}
%-----
where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.

Thus far, we have considered only a single source of information to fit this model, and therefore the prior distribution \(p\left(\theta_M \right)\) from Eq. \ref{eq:ex1_bayes} is uninformative.
As a secondary source of information, we will consider an experiment relating the population growth rate of the plant to manipulations to the precipitation regime, with results (but no raw data) available from the literature (Fig. \ref{fig:ex1_sampling}b). 
Furthermore, no information is available regarding the temperature regime for the experiment.
Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
This concept gives us a reasonable model to fit a scaling function for our sub-model (Appendix S1).
If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed (where the \(S\) subscript indicates information pertaining to the submodel), then for an observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
\begin{equation}
	\psi_{S,i} = \int_0^\infty N \left(X_{S,i}, \sigma_{S,i} \right)
\end{equation}
where \(N\) is the normal density function.
We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
%-----
\begin{equation}
\label{eq:ex1_thetas}
	p\left (\theta_S \mid \psi_S,P_S \right ) \propto 
	p \left(\psi_S \mid \theta_S, P_S \right)
	p \left(\theta_S \right) \\
\end{equation}
%-----

Although the two datasets were collected at considerably different scales, we have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation). 
The scaling treats the fundamental niche as the only driver of species distribution, and only considers a single dimension of the niche.
As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
However, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.

We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
%-----
\begin{equation}
	\label{eq:ex1_integrated}
	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S, P_S)}^\text{integrated posterior}
	\propto
	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
\end{equation}
%-----
As before, the metamodel \(\theta_M\) can be used to predict probability of occurrence \((\psi_I)\).
However, these predictions now reflect the presence-absence data \(X_M\) as well as the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).

%==================
\begin{figure}[bt]
	\includegraphics{figs/ex1_precip.pdf}
	\caption{Comparison of the naive model, mechanistic sub-model, and the integrated model showing the probability of presence (\(\psi\)) as a function of precipitation.
	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals.
	The shaded region shows the calibration range for the naive model.
	(a) Naive model, using only presence-absence data. Uncertainty increases dramatically when attempting to project beyond the scope of the source data.
	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
	}
	\label{fig:ex1_precip}
\end{figure}
%==================
